Generic Properties of D-optimal Designs in Infinite Dimensional Bayesian Linear Inverse Problems

### 2023-04-02 latex/intro.tex
getting closegit add latex/vanishing.tex !

When no model error is present, the Lagrange multipliers problem boils
down to a nonlinear eigenvalue problem.
%% The operator for which eigenvectors and eigenvalues are sought is a
%% sum of two operators. The first is the prior covariance. The second
%% is an outer product of the observations (see section
%% \ref{section:vanishing} for details and exact statement).
This structure helps me characterize D-optimal designs in Theorem
\ref{thm:char}. The key idea is that a D-optimal design reduces
uncertainties where they are highest first.

\begin{restatable}[D-optimal designs with vanishing model error]{theorem}{main}\label{thm:char}
  Let:
  \begin{enumerate}
  \item $\fwd:\hilp \to \hilo$ a linear forward operator,
    $\prcov:\hilp \to \hilp$ prior covariance operator, $\obs: \hilo
    \to \mathbb{R}^m$ observation operator, where $m \in \mathbb{N}$
    is the number of observations taken.
    \item $\sigma^2 \in \mathbb{R}_{+}$ observation noise variance,
      $\data = \obs \fwd \param + \eps$, where $\eps \in \mathbb{R}^m$
      is iid $\mathcal{N}(0, \sigma^2)$ noise, $\pr \sim
      \mathcal{N}(0, \prcov)$ prior measure, $\post$ the posterior
      measure.
  \item A D-optimality utility function
    \cite{AlexanderianGloorGhattas14}:
    \begin{align*}
      \begin{split}
        \tar(\obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
        % 
        % 
        % 
        &= \frac12 \log \det ( I + \sigma^{-2} \prcov^{1/2} \fwd ^*
        \obs^* \obs \fwd \prcov^{1/2}).
    \end{split}
  \end{align*}
  \item A D-optimal design operator $\obs$:
    $$
    \obs = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs)
    $$ 
  \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of $\fwd\prcov\fwd^*$
    in decreasing order of magnitude.
  %% \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$.
  \end{enumerate}

  Then:
  \begin{enumerate}
  \item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously
    diagonalizable.
  \item $k := \rank \obs^*\obs \leq m$.
  \item     
    \begin{equation*}
      \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). %= \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
    \end{equation*}
  %% \item 
  %%   \begin{equation*}
  %%     k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
  %%     \lambda_j^{-1} \right \}.
  %%   \end{equation*}
  \item
    \begin{equation*}
        \eta_i = \begin{cases}
          \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & 1 \leq i \leq k \\
          0 & i > k 
        \end{cases}.
    \end{equation*}
  \end{enumerate}
\end{restatable}

@@

In section \ref{section:prelim} I present a more abstract and general
formulation of the inverse problem I consider. But, for the purpose of
illustration, I present clusterization via a toy model --- the 1D heat
equation in $[0,\pi]$ with a homogeneous Dirichlet boundary condition.

>>>>

When no model error is present, the Lagrange multipliers problem boils
down to a nonlinear eigenvalue problem.
%% The operator for which eigenvectors and eigenvalues are sought is a
%% sum of two operators. The first is the prior covariance. The second
%% is an outer product of the observations (see section
%% \ref{section:vanishing} for details and exact statement).
This structure helps me characterize D-optimal designs. The key idea
is that a D-optimal design reduces uncertainties where they are
highest first. This is the main conclusion of Theorem \ref{thm:char}
stated below.

\begin{restatable}[D-optimal designs with vanishing model error]{theorem}{char}\label{thm:char}
%% \begin{theorem}[D-optimal designs with vanishing model error]\label{thm:char}
  Let:
  \begin{itemize}
  \item $\hilp, \hilo$ Hilbert spaces.
  \item $\fwd:\hilp \to \hilo$ a linear compact operator.
  \item $\pr \sim \mathcal{N}(0, \prcov)$ prior Gaussian measure on $\hilp$,
    where $\prcov:\hilp \to \hilp$ is the prior covariance operator.
  \item $\obs: \hilo \to \mathbb{R}^m$ observation operator, where $m
    \in \mathbb{N}$ is the number of observations taken. 
  \item $\sigma^2 \in \mathbb{R}_{+}$ observation noise variance,
    $\data = \obs \fwd \param + \eps$, where $\eps \in \mathbb{R}^m$
    is iid $\mathcal{N}(0, \sigma^2)$ noise.
  \item $\post$ the posterior measure, with covariance $\postcov$.
  \item A D-optimality utility function
    \cite{AlexanderianGloorGhattas14}:
    \begin{align*}
      \begin{split}
        \tar(\obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
        % 
        % 
        % 
        &= \frac12 \log \det ( I + \sigma^{-2} \prcov^{1/2} \fwd ^*
        \obs^* \obs \fwd \prcov^{1/2}).
      \end{split}
    \end{align*}
  \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of $\fwd\prcov\fwd^*$
    in decreasing order of magnitude.
  %% \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$.
  \end{itemize}

  If $\obs$ is a D-optimal design operator
    \begin{equation*}
      \obs = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs), 
    \end{equation*}
    then:
  \begin{enumerate}
  \item  $\tr{\obs^*\obs} = m$.
  \item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously
    diagonalizable.
  \item $k := \rank \obs^*\obs \leq m$ and
    \begin{equation*}
      \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). %= \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
    \end{equation*}
  %% \item 
  %%   \begin{equation*}
  %%     k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
  %%     \lambda_j^{-1} \right \}.
  %%   \end{equation*}
  \item
    \begin{equation*}
        \eta_i = \begin{cases}
          \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & 1 \leq i \leq k \\
          0 & i > k 
        \end{cases}.
    \end{equation*}
  \item The covariance of the pushforwad $\fwd_{*} \post$ is $\left (
    (\fwd \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs \right )^{-1}$
    and has eigenvalues
    \begin{equation*}
      \theta_i =
      \begin{cases}
        \left(\frac{\sum_{j=1}^k \lambda_j^{-1} + \sigma^{-2}m}{k} \right )^{-1} & i \leq k \\
        \lambda_i &  i > k 
      \end{cases}
    \end{equation*}
  \end{enumerate}
%\end{theorem}
\end{restatable}

%% \begin{restatable*}[Flat posterior covariance]{corollary}{flatpost}\label{cor:equal eigs}
%%  %\begin{corollary}\label{cor:eigenvalues}
%%   If $\obs$ is D-optimal and $k := \rank \obs$, then:
%%   \begin{enumerate}
%%     \item The covariance of the pushforwad $\fwd_{*} \post$ is $\left ( (\fwd
%%       \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs \right )^{-1}$.
%%     \item The eigenvalues of said operator are
%%       \begin{equation}
%%         \theta_i =
%%         \begin{cases}
%%           \left(\frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} + \sigma^{-2}\frac{m}{k} \right )^{-1} & i \leq k \\
%%           \lambda_i &  i > k 
%%         \end{cases}
%%       \end{equation}
%%     \item The eigenvectors of said operator are the eigenvectors of
%%       $\fwd \prcov \fwd^*$
%%   \end{enumerate}
%% \end{restatable*}%\end{corollary}

%% \begin{restatable*}[Flat posterior covariance]{corollary}{flatpost}\label{cor:equal eigs}
%%   Consider a D-optimal design $\obs$ for a Bayesian linear inverse
%%   problem where $\fwd$ is the forward operator and $\pr$ is the
%%   prior. Then the eigenvectors of the covariance of the pushforward of
%%   the posterior $\fwd_{*} \post$ equal the eigenvectors of the
%%   covariance of the pushforward of the prior. Moreover, the
%%   eigenvalues are constant for eigenvectors $1 \leq i \leq k$ where $k
%%   =\rank \obs$.
%% \end{restatable*}

@@

In section \ref{section:prelim}, I present a more abstract and general
formulation of the inverse problem I consider. But, for the purpose of
illustration, I present clusterization via a toy model --- the 1D heat
equation in $[0,\pi]$ with a homogeneous Dirichlet boundary condition.

### 2023-04-02 latex/prelims.tex
getting closegit add latex/vanishing.tex !

\subsection{Overview}
In this section we define notations that will be used throughout the
paper. The theoretical foundations for inverse problems over function spaces
can be found in \cite{Stuart10}.

\subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
Let $\hilp$ and $\hilo$ be separable Hilbert spaces (the subscripts p
and o are for ``parameter'' and ``observation'', respectfully). We
denote $\hilo^*$ the Hilbert space of all linear functionals on
$\hilo$. Consider $\fwd: \hilp \to \hilo$, a linear operator (the
``forward operator''). We are interested in forward operators that are
strongly smoothing (have fast decaying modes --- the heat operator
from section \ref{subsec:example} is a prime example). Our goal is to
infer $\param \in \hilp$ --- some parameter of the dynamics --- given
noisy observations of $\fwd \param$. We take a Gaussian prior $\pr$
for $\param$. Hence, $\param \sim \pr = \normal(\prmean ,\prcov)$ with
some appropriate covariance operator $\prcov$ \cite{Stuart10}. It is
important to note that $\fwd \prcov \fwd^*$ is the prior covariance in
$\hilo$ \cite{Stuart10}. As such, it is assumed invertible --- an
assumption we rely on below. Throughout this paper, we denote $m$ the
number of observations we are allowed to take. Observations are taken
via the operator $\obs \in ( \hilo^* )^m$. In an analogy with linear
algebra, where row vectors are thought of as linear functionals, we
think of $\obs$ as having ``rows'' $\meas_j, j=1,\dots,m$:
\begin{equation}\label{eq:O}
  \obs = (\meas_1,\dots, \meas_m)^t, \meas_j \in \hilo^*, j = 1,\dots,m.
\end{equation}
This way, for $u \in \hilo$ we have $\obs u = (\meas_1(u), \dots,
\meas_m(u) )^t \in \R^m$.
%% but we may drop the parentheses and write $\meas_j u = \meas_j(u)$.
%% A few observations regarding $\obs$ are in order.
%% First, it is good to keep in mind that $( \hilo^* )^m$ is a Hilbert
%% space with norm $\| \obs \| = \sum_{j=1}^m \|\meas_j\|$.
For $u \in \hilo$ and $v\in \R^m$:
\begin{align*}
  \big (\obs^*v \big ) (u) &= \langle v, \obs u \rangle_{\R^m} = \sum_{j=1}^m  v_j \meas_j(u)
  = v^t \left ( \obs u \right ) = (v^t \obs) (u),
\end{align*}
and thus:
\begin{align}\label{eq:obs*}
  \obs^*v &= \sum_{j=1}^m v_j \meas_j = v^t \obs.
\end{align}

@@

Each observation is a linear functional, chosen from some subset of
$\hilo^{*}$. Data is acquired via noisy observations
\begin{align}\label{eq:inverse problem}
  \data := \obs (\fwd \param + \eps') + \eps = \obs \fwd \param + \obs \eps' + \eps,
\end{align}
with $\data \in \R^m$ and $\eps, \eps'$ defined next. We consider two
types of error terms. First, there is spatially correlated model error
$\eps' \sim \normal(0,\modcov)$, modeled as a centered Gaussian
measure on $\hilo$ with covariance operator $\modcov$. Then there is
observation error $\eps \sim \normal(0, \sigma^2 I_m)$, with $I_m$ the
$m \times m$ identity. Both error terms and the prior are assumed
independent of each other. The observation and forward operators are
typically merged $\tmp := \obs \fwd$
\cite{AlexanderianGloorGhattas14}. Since we need to separate the
observation operator $\obs$ from the forward operator $\fwd$ in order
to carry out the analysis in the following sections, we explicitly
separate $\fwd$ and $\obs$ in \eqref{eq:inverse problem}. We now turn
to understand the terms present in \eqref{eq:inverse problem}.

\subsubsection{Dynamics and Observation Operators}\label{subsec:dynamics}
Rows of $\obs$ are linear functionals and we refer to these as {\it
  observation vector}s. Note that we cannot choose every observation
vector we want. For example, we may be restricted only to pointwise
evaluations of $\fwd \param$ by the sensors at our disposal, in which
case measuring, e.g., the mean $\ell(u) = \int_{\Omega}u$ of a
function is not possible. Thus, it is best to think of $(\obs \fwd
\param)_j$ as approximations to pointwise evaluations. For example, if
$\hilo = L^2([-\pi, \pi])$, consider $\ell_1(x) := (2\pi
\eps)^{-\frac{1}{2}}\exp(-\frac{x^2}{2\eps^2})$ and $\ell_2(x)
:=\eps^{-1}\mathbf{1}_{[-\frac{\eps}{2}, \frac{\eps}{2}]}$. Both
$\ell_1$ and $\ell_2$ approximate a point measurement at $0$ for small
$\eps$ and, crucially, $\ell_1, \ell_2 \in \hilo^*$. It is important
to note that although $\|\ell\|_1 = \|\ell_2\|_1=1$, unfortunately
$\|\ell_1\|_2^2 = (4\pi \eps)^{-\frac{1}{2}}$ and $\|\ell_2\|_2^2 =
\eps^{-1}$.

\subsubsection{Error Terms}
Considering the sum of the error terms, it is easy to see that
$\bar{\eps} := \obs \eps' + \eps \in \R^m$. It is a centered Gaussian
random vector and its covariance matrix is
\begin{align}\label{eq:Sigma}
  \begin{split}
    \Sigma(\obs) :&= \mathbb{E}[ (\obs \eps' + \eps)  (\obs \eps' + \eps)^t ] 
    % 
    % 
    = \obs \modcov \obs^* + \sigma^2I , 
  \end{split}
\end{align}
where
\begin{equation}\label{eq:modcov explained}
  [\obs \modcov \obs^*]_{ij} = e_i^t \obs \modcov \obs^* e_j = \meas_i (\modcov \meas_j),
\end{equation}
with $e_j \in \R^m$ the $j$th standard basis vector. The first
equality in \eqref{eq:modcov explained} holds by definition and the
second by \eqref{eq:obs*}. The explicit dependence on $\obs$ will be
mostly dropped for notational convenience, so $\Sigma(\obs) =
\Sigma$. Thus, for a fixed $\obs$ (i.e.\ a fixed set of observations)
we can equivalently write $\data = \obs \fwd \param + \bar{\eps}$ with
$\bar{\eps} \sim \normal(0,\Sigma)$. Taking $\modcov = 0$ is common
practice \cite{Tarantola05,Kaipio2006,Vogel02}. This reduces to
the case where we take iid observations. Then $\Sigma = \sigma^2I$ is
simply a scalar matrix that does not depend on $\obs$. Note that
taking an error model with a non-scalar covariance as we do here
allows us to consider model error (modeled by $\modcov$) as well as
observation error (modeled by $\sigma^2$). For example, say we believe
our forward model does not capture some small scale phenomenon.  Then
we may express this belief by saying $\tru = \fwd + \err$, with $\fwd$
depending on $\param$ and $\err$ depending on $\sspar$, with $\param
\perp \sspar$. We do not know much about this effect but it is
reasonable to assume it changes continuously in our domain. We (may
choose to) model it as $\normal (0, \modcov)$ and take $\modcov$ to
reflect the spatial (or other) variability we imagine $\err \sspar$
has. Such small scale phenomenon can arise as a modeling issue, where
we might not model the system in its entirety. It can also arise from
a numerical source, where our discretization of the system is not fine
enough to capture all small scale phenomena. In section
\ref{section:non vanishing}, we will see that assuming some
correlation in the error mitigates the clusterization phenomenon, as
reported in the literature \cite{Ucinski05}.

@@

\subsection{D-Optimal Designs in Infinite Dimensions}\label{subsec:D optimal design} 
A D-optimal design maximizes expected KL divergence between posterior
and prior. It is useful to recall the definition of KL divergence for
an arbitrary prior measure:
$$
KL(\post||\pr) = \int \log \frac{\der \post}{\der \pr}(\param) \der \post(\param).
$$

The meaning of D-optimal design in infinite-dimensional Hilbert spaces
was investigated in \cite{AlexanderianGloorGhattas14}. There, the
authors make assumptions that amount to $\Sigma=I$ (implied by
$\modcov = 0,\sigma^2=1$), but we choose not to take these
simplification here. This is because $\Sigma$ can determine ``how
much'' clusterization we see. As stated
\cite[pp. 681]{AlexanderianGloorGhattas14}, the results hold for more
general covariance matrices. The conclusion is that in
infinite-dimensions, a D-optimal design is well-defined as maximizing
the expected KL divergence between posterior and prior. The main
result is summarized, using our notation, in the following theorem:
\begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]\label{thm:d optimality}
  Let $\pr = \normal(\prmean,\prcov)$ be a Gaussian prior on $\hilp$
  and let $\post = \normal(\postmean,\postcov)$ the posterior measure
  on $\hilo$ for the Bayesian linear inverse problem $\data = \obs
  \fwd\param + \obs \eps' + \eps$ discussed above. Then
  \begin{align}\label{eq:objective}
    \begin{split}
      \tar( \obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
      % 
      % 
      % 
      &= \frac12 \log \det 
      ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}).
    \end{split}
  \end{align}
\end{theorem}
\begin{definition}\label{def:d optimality}
  A design $\obs^{\star}$ is said to be D-optimal if $\obs^{\star} = \argmax_{\obs} \tar(\obs)$.
\end{definition}

>>>>

\subsection{Overview}
In this section I will define notations that will be used throughout
the paper. The theoretical foundations for inverse problems over
function spaces can be found in \cite{Stuart10}. Most of the notations
are summarized in the conditions of Theorem \ref{thm:char}.

\subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
Let $\hilp$ and $\hilo$ be separable Hilbert spaces (the subscripts p
and o are for ``parameter'' and ``observation'', respectfully). Denote
$\hilo^* \simeq \hilo$ the Hilbert space of all linear functionals on
$\hilo$. Consider $\fwd: \hilp \to \hilo$, a linear operator (the
``forward operator''). We are interested in forward operators that are
strongly smoothing (have fast decaying modes --- the heat operator
from section \ref{subsec:example} is a prime example). Our goal is to
infer $\param \in \hilp$ --- some parameter of the dynamics --- given
noisy observations of $\fwd \param$. We take a Gaussian prior $\pr$
for $\param$. Hence, $\param \sim \pr = \normal(\prmean ,\prcov)$ with
some appropriate covariance operator $\prcov$ \cite{Stuart10}. It is
important to note that $\fwd \prcov \fwd^*$ is the prior covariance in
$\hilo$ \cite{Stuart10}. As such, it is assumed invertible --- an
assumption we rely on below. Throughout this paper, we denote $m$ the
number of observations we are allowed to take. Observations are taken
via the operator $\obs \in ( \hilo^* )^m$. In an analogy with linear
algebra, where row vectors are thought of as linear functionals, we
think of $\obs$ as having ``rows'' $\meas_j, j=1,\dots,m$:
\begin{equation}\label{eq:O}
  \obs = (\meas_1,\dots, \meas_m)^t, \meas_j \in \hilo^*, j = 1,\dots,m.
\end{equation}
This way, for $u \in \hilo$ we have $\obs u = (\meas_1(u), \dots,
\meas_m(u) )^t \in \R^m$.
%% but we may drop the parentheses and write $\meas_j u = \meas_j(u)$.
%% A few observations regarding $\obs$ are in order.
%% First, it is good to keep in mind that $( \hilo^* )^m$ is a Hilbert
%% space with norm $\| \obs \| = \sum_{j=1}^m \|\meas_j\|$.
For $u \in \hilo$ and $v\in \R^m$:
\begin{align*}
  \big (\obs^*v \big ) (u) &= \langle v, \obs u \rangle_{\R^m} = \sum_{j=1}^m  v_j \meas_j(u)
  = v^t \left ( \obs u \right ) = (v^t \obs) (u),
\end{align*}
and thus:
\begin{align}\label{eq:obs*}
  \obs^*v &= \sum_{j=1}^m v_j \meas_j = v^t \obs.
\end{align}

@@

Each observation is a linear functional, chosen from some subset of
$\hilo^{*}$. Data is acquired via noisy observations
\begin{align}\label{eq:inverse problem}
  \data := \obs (\fwd \param + \eps') + \eps = \obs \fwd \param + \obs \eps' + \eps,
\end{align}
with $\data \in \R^m$ and $\eps, \eps'$ defined next. We consider two
types of error terms. First, there is spatially correlated model error
$\eps' \sim \normal(0,\modcov)$, modeled as a centered Gaussian
measure on $\hilo$ with covariance operator $\modcov$. Then there is
observation error $\eps \sim \normal(0, \sigma^2 I_m)$, with $I_m$ the
$m \times m$ identity. Both error terms and the prior are assumed
independent of each other. In the literature, it is common for the
observation and forward operators to be merged $\tmp := \obs \fwd$
\cite{AlexanderianGloorGhattas14}. Since we need to separate the
observation operator $\obs$ from the forward operator $\fwd$ in order
to carry out the analysis in the following sections, we explicitly
separate $\fwd$ and $\obs$ in \eqref{eq:inverse problem}. We now turn
to understand the terms present in \eqref{eq:inverse problem}.

\subsubsection{Dynamics and Observation Operators}\label{subsec:dynamics}
Rows of $\obs$ are linear functionals and we refer to these as {\it
  observation vector}s. Note that we cannot choose every observation
vector we want. For example, we may be restricted only to pointwise
evaluations of $\fwd \param$ by the sensors at our disposal, in which
case measuring, e.g., the mean $\ell(u) = \int_{\Omega}u$ of a
function is not possible. Thus, it is best to think of $(\obs \fwd
\param)_j$ as approximations to pointwise evaluations. For example, if
$\hilo = L^2([-\pi, \pi])$, consider $\ell_1(x) := (2\pi
\eps)^{-\frac{1}{2}}\exp(-\frac{x^2}{2\eps^2})$ and $\ell_2(x)
:=\eps^{-1}\mathbf{1}_{[-\frac{\eps}{2}, \frac{\eps}{2}]}$. Both
$\ell_1$ and $\ell_2$ approximate a point measurement at $0$ for small
$\eps$ and, crucially, $\ell_1, \ell_2 \in \hilo^*$.
%% It is important to note that although $\|\ell\|_1 =
%% \|\ell_2\|_1=1$, unfortunately $\|\ell_1\|_2^2 = (4\pi
%% \eps)^{-\frac{1}{2}}$ and $\|\ell_2\|_2^2 = \eps^{-1}$.

\subsubsection{Error Terms}
Considering the sum of the error terms, it is easy to verify that
$\bar{\eps} := \obs \eps' + \eps \in \R^m$. It is a centered Gaussian
random vector and its covariance matrix is
\begin{align}\label{eq:Sigma}
  \begin{split}
    \Sigma(\obs) :&= \mathbb{E}[ (\obs \eps' + \eps)  (\obs \eps' + \eps)^t ] 
    % 
    % 
    = \obs \modcov \obs^* + \sigma^2I , 
  \end{split}
\end{align}
where
\begin{equation}\label{eq:modcov explained}
  [\obs \modcov \obs^*]_{ij} = e_i^t \obs \modcov \obs^* e_j = \meas_i (\modcov \meas_j),
\end{equation}
with $e_j \in \R^m$ the $j$th standard basis vector. The first
equality in \eqref{eq:modcov explained} holds by definition and the
second by \eqref{eq:obs*}. The explicit dependence on $\obs$ will be
mostly dropped for notational convenience, so $\Sigma(\obs) =
\Sigma$. Thus, for a fixed $\obs$ (i.e.\ a fixed set of observations)
we can equivalently write $\data = \obs \fwd \param + \bar{\eps}$ with
$\bar{\eps} \sim \normal(0,\Sigma)$. Taking $\modcov = 0$ is common
practice \cite{Tarantola05,Kaipio2006,Vogel02}. This reduces to
the case where we take iid observations. Then $\Sigma = \sigma^2I$ is
simply a scalar matrix that does not depend on $\obs$. Note that
taking an error model with a non-scalar covariance as we do here
allows us to consider model error (modeled by $\modcov$) as well as
observation error (modeled by $\sigma^2$). For example, say we believe
our forward model does not capture some small scale phenomenon.  Then
we may express this belief by saying $\tru = \fwd + \err$, with $\fwd$
depending on $\param$ and $\err$ depending on $\sspar$, with $\param
\perp \sspar$. We do not know much about this effect but it is
reasonable to assume it changes continuously in our domain. We (may
choose to) model it as $\normal (0, \modcov)$ and take $\modcov$ to
reflect the spatial (or other) variability we imagine $\err \sspar$
has. Such small scale phenomenon can arise as a modeling issue, where
we might not model the system in its entirety. It can also arise from
a numerical source, where our discretization of the system is not fine
enough to capture all small scale phenomena.
%%  In section \ref{section:non vanishing}, we will see that assuming
%% some correlation in the error mitigates the clusterization
%% phenomenon, as reported in the literature \cite{Ucinski05}.

@@

\subsection{D-Optimal Designs in Infinite Dimensions}\label{subsec:D optimal design} 
A D-optimal design maximizes expected KL divergence between posterior
and prior measures, denoted $\post$ and $\pr$, respecively. It is
useful to recall the definition of KL divergence for an arbitrary
prior measure:
$$
KL(\post||\pr) = \int \log \frac{\der \post}{\der \pr}(\param) \der \post(\param).
$$

The meaning of D-optimal design in infinite-dimensional Hilbert spaces
was investigated in \cite{AlexanderianGloorGhattas14}. There, the
authors make assumptions that amount to $\Sigma=I$ (implied by
$\modcov = 0,\sigma^2=1$), but we choose not to take these
simplification here. This is because $\Sigma$ can determine ``how
much'' clusterization we see. As stated
\cite[pp. 681]{AlexanderianGloorGhattas14}, the results hold for more
general covariance matrices. The conclusion is that in
infinite-dimensions, a D-optimal design is well-defined as maximizing
the expected KL divergence between posterior and prior. The main
result is summarized, using our notation, in the following theorem:
\begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]\label{thm:d optimality}
  Let $\pr = \normal(\prmean,\prcov)$ be a Gaussian prior on $\hilp$
  and let $\post = \normal(\postmean,\postcov)$ the posterior measure
  on $\hilp$ for the Bayesian linear inverse problem $\data = \obs
  \fwd\param + \obs \eps' + \eps$ discussed above. Then
  \begin{align}\label{eq:objective}
    \begin{split}
      \tar( \obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
      % 
      % 
      % 
      &= \frac12 \log \det 
      ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}).
    \end{split}
  \end{align}
\end{theorem}
\begin{definition}\label{def:d optimality}
  A design $\obs^{\star}$ is said to be D-optimal if $\obs^{\star} = \argmax_{\obs} \tar(\obs)$.
\end{definition}
